### 2023-09-18 doc/_discussion.qmd
finish first draft of discussion

In this paper, we present a computational pipeline to analyze lineage-tracking
time-series data for massive-parallel competition assays. More specifically, we
fit a Bayesian model to infer the fitness of multiple genotypes relative to a
reference [@kinsler2020;@ascensao2023]. The proposed model accounts for multiple
sources of uncertainty with proper error propagation intrinsic to Bayesian
methods. To scale the inference pipeline to large datasets with $> 10,000$
barcodes, we use the ADVI algorithm [@kucukelbir2016] to fit a variational
posterior distribution. The main difference between our method and previous
inference pipelines, such as @li2023, is that the present analysis provides
interpretable errors on the inferred fitness values. The reported uncertainty
intervals---known as credible regions---can be adequately interpreted as
capturing the corresponding probability mass of finding the true value of the
parameter given the model, the prior information, and the data. Furthermore,
minor modifications to the structure of the statistical model allow for the
analysis of different experimental designs, such as growth-dilution cycles in
different environments, joint analysis of multiple experimental replicates of
the same experiment via hierarchical models, and a hierarchical model for
multiple barcodes mapping to equivalent genotypes.

@@

>>>>

In this paper, we present a computational pipeline to analyze lineage-tracking
time-series data for massive-parallel competition assays. More specifically, we
fit a Bayesian model to infer the fitness of multiple genotypes relative to a
reference [@kinsler2020;@ascensao2023]. The proposed model accounts for multiple
sources of uncertainty with proper error propagation intrinsic to Bayesian
methods. To scale the inference pipeline to large datasets with $> 10,000$
barcodes, we use the ADVI algorithm [@kucukelbir2016] to fit a variational
posterior distribution. The main difference between our method and previous
inference pipelines, such as @li2023, is that the present analysis provides
interpretable errors on the inferred fitness values. The reported uncertainty
intervals---known as credible regions---can be formally interpreted as capturing
the corresponding probability mass of finding the true value of the parameter
given the model, the prior information, and the data. Furthermore, minor
modifications to the structure of the statistical model presented in this work
allow for the analysis of different experimental designs, such as
growth-dilution cycles in different environments, joint analysis of multiple
experimental replicates of the same experiment via hierarchical models, and a
hierarchical model for multiple barcodes mapping to equivalent genotypes.

@@

It is important to highlight some of the consequences of the general
experimental design and the implicit assumptions within the proposed statistical
model to analyze the resulting data. First, the composition of the population is
such that the initial fraction of the population occupied by the barcoded
genotypes is small---usually >90\% of the initial population is the non-labeled
reference strain. This constraint is important as the fitness model used to fit
the time series data assumes that the tracked frequencies are $\ll 1$. Second,
when computing log frequency ratios, we can run into the issue of dividing by
zero. This is a common problem when dealing with molecular count data
[@lovell2020]. Our model gets around this issue by assuming that the frequency
of any barcode cannot be, but still can get arbitrarily close to, zero.
Therefore, we implicitly assume that no lineage goes extinct during the
experiment. Moreover, the statistical model directly accounts for the
uncertainty associated with having zero barcode counts, increasing the
corresponding uncertainty. Third, the models presented in this paper require the
existence of a labeled sub-population of barcoded reference strains. These
barcodes help determine the fitness baseline, as every fitness is quantified
with respect to this reference genotype. This experimental design constraint
facilitates the inference of the population mean fitness since most of the
culture---the unlabeled reference genotype---is not tracked. Finally, the
presented statistical model assumes that relative fitness is solely a constant
of the environment and the genotype. Therefore, the model is not designed to
analyze data with time-varying or frequency-dependent fitness values.
